---
title: "Penguin Size Comparisons"
author: "Alison Sells"
date: "11/14/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(palmerpenguins)
library(effectsize)
library(broom)
library(ggbeeswarm)
library(car)
```

# Part 1. T-tests

Mass comparison between male and female chinstrap penguins

Exploring of data

```{r}
chinstraps<- penguins %>% 
  filter(species== "Chinstrap")
```

#Histogram
```{r}
ggplot(data = chinstraps, aes(x= body_mass_g))+
  geom_histogram(bins= 10) +
  facet_wrap(~sex)
```
# QQ plot
```{r}
ggplot(data = chinstraps, aes (sample= body_mass_g))+
geom_qq()+
  facet_wrap(~sex)
```


## Get Descriptive Statistics

```{r}
chinstraps %>% 
  group_by(sex) %>% 
  summarize(
    mean_mass= mean(body_mass_g),
    median_mass= median(body_mass_g),
    sd_mass= sd(body_mass_g),
    sample_size = n()
  )
  
```
Get vectors of body mass 4 female ans male chinstrap penguins sep.


```{r}
chinstraps_f<- chinstraps %>% 
  filter(sex== "female") %>% 
  pull(body_mass_g)

chinstraps_m<- chinstraps %>% 
  filter(sex== "male") %>% 
  pull(body_mass_g)


```

## T-tests, a couple of different ways

1. is there a sig. difference in body mass between male and female chinstrap penguins
```{r}
chinstrap_ttest<- t.test(chinstraps_f, chinstraps_m)
chinstrap_ttest


cohens_d(chinstraps_f, chinstraps_m)
```

Are male penguins sig. larger than females (1 sided)

```{r}
t.test(chinstraps_m, chinstraps_f, alternative = "greater")
```
## One way ANOVA

- single varialbe that has discrete grouos within it
another contineud mesaured variable
- extrension of the 2 sided t test
## 1. Table of counts by species and sex

```{r}
penguins_counts <- penguins %>% 
  count(species, sex)

penguins_counts
```


Create a subset of only females

```{r}
female_chinstraps<- penguins %>% 
  filter(sex== "female")

ggplot(data = female_chinstraps, aes (x= flipper_length_mm ))+
geom_histogram(aes(fill= species))+
  facet_wrap(~species, scales = "free")
  

ggplot(data = female_chinstraps, aes (sample= flipper_length_mm ))+
geom_qq(aes(color= species))+
  facet_wrap(~species, scales = "free")
```
## summary stats
cehck for equal vairances

```{r}
penguin_f_summary <- female_chinstraps %>% 
  group_by(species) %>% 
  summarize(
    mean_flipper_length= mean(flipper_length_mm),
    sd_flipper_length= sd(flipper_length_mm),
    sample_size= n(),
    se_flipper_length = sd(flipper_length_mm)/ sqrt(n()),
    var_flipper_length= var(flipper_length_mm)
  )



leveneTest(flipper_length_mm~species, data= female_chinstraps)
```

### Beeswarm Plot
with mean and +- 1sd

```{r}
ggplot()+ geom_beeswarm(data= female_chinstraps,
                        aes(x= species,y= flipper_length_mm),
color= "gray70",
size= 3,
alpha= 0.5)+
geom_point(data = penguin_f_summary, 
           aes(x= species,y= mean_flipper_length),
           size= 4)+
  geom_errorbar(data = penguin_f_summary,
                aes(x= species, ymax=  mean_flipper_length + sd_flipper_length ,
                 ymin= mean_flipper_length - sd_flipper_length),
width= 0.3) +
  theme_light()

```

## 1 way ANOVA


```{r}
penguin_aov<- aov(flipper_length_mm~ species, data= female_chinstraps)
summary(penguin_aov)


## Post Hoc Test

TukeyHSD(penguin_aov)
```
If very unbalanced design used type 3 ANOVA with car package (create a linear model)
```{r}
penguin_lm <- lm(flipper_length_mm~ species, data= female_chinstraps) 
penguin_aov_t3 <- Anova(penguin_lm, type= "III")
penguin_aov_t3
```


